---
title: "chord.chain"
author: "LA"
date: "2022-10-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## MARKOV CHAIN - CHORD PROGRESSION
Music can be very complex. A song can consist of a series of chords, harmonies, melodies, and vocals.
Chord progressions are often written in roman numerals, where upper case numerals are for major chords, and lower case numerals are for minor chords. 
I want to look at the chord progression sequences of songs, as well as the distribution of these chord progressions using markov chains Chord progressions.

Chords are labeled 1-7 (I ii iii IV V vi vii) . 

I will analyze two songs by one artist and compare the probability matrices with two songs by a different artist. They should be different
```{r}
library(markovchain)
library(dplyr)

# SONGS AND THEIR KEY CHORD PROGRESSIONS (more songs will be added)

chords <- c("I", "ii", "iii", "IV", "V", "vi", "vii")

#chords <- c("I", "i", "II", "ii", "III", "iii", "IV","IV7", "iv", "V","v","VI", "vi", "VII", "vii", "vii7")

#chords <- c("I", "ii", "iii", "IV", "IV7" ,"V", "vi", "vii", "vii7")

BadRomance <- c("I", "Vii", "vi", "IV", "V", "I")
Dreams <- c("vi","V", "IV", "V", "vi", "V", "IV")
Hallelujah <- c("V","iii", "V", "iii", "I", "ii", "V")
ForeverYoung <- c("vi", "I", "vi", "I", "IV", "vi", "V", "IV", "vi", "V", "IV", "vi","V", "IV", "vi", "V", "i","VI")
SevenWonders <- c("IV","I","vi","V","IV","I","vi","V","IV","V","vi","IV",
                  "I", "V", "I", "V", "I", "V", "IV", "V")

#FINDING PROBABILITIES 

songs <- list(BadRomance,Dreams,Hallelujah)
songs.MC <- markovchainFit(data=songs)

songs.MC

##TRANSITION/MARKOV CHAIN APPLICATION

chords <- c("I", "ii", "iii", "IV", "V", "vi", "vii")
chord.transition <- matrix(c(0,0,0,0.5,0.5,0,0,
                             0,0,1,0,0,0,0,
                             0.1666667,0.3333333,0,0,0,0.3333333,0.1666667,
                             0,0,0,0,0,0,1,
                             0,0,1,0,0,0,0,
                             0.5,0,0.5,0,0,0,0,
                             0,0.3333333,0.6666667,0,0,0,0), nrow=7, ncol=7, byrow=TRUE)

MC.chords <- new("markovchain", states=chords, byrow=TRUE,transitionMatrix=chord.transition, name="Chord Progression")

class(MC.chords)
MC.chords^3

PWF <- markovchainFit(PlayingWithFire,sanitize=TRUE, possibleStates = possible.chords)

LG <- markovchainFit(LovesickGirls,sanitize=TRUE, possibleStates = possible.chords)

PWF
LG



#Creating Matrix

Blackpink <- c(PlayingWithFire,LovesickGirls)

BP.chordcounts <- createSequenceMatrix(Blackpink,toRowProbs=FALSE,sanitize=FALSE, possibleStates = possible.chords)

BP.chordcounts

BP.chordprob <- createSequenceMatrix(Blackpink,toRowProbs=TRUE,sanitize=FALSE, possibleStates = possible.chords)

BP.chordprob

#Transition Probability Matrix 

BP.TM <- markovchainFit(Blackpink,sanitize=TRUE, possibleStates = possible.chords)

BP.TM

```




I will be comparing two songs from the girl-group BLACKPINK, and two songs from the band Fleetwood Mac
```{r}

library(markovchain)

#Minor chords are lowercase
#Major chords are uppercase

possible.chords <- c("I","i","II", "ii", "III","iii", "IV","iv", "V", "v","VI", "vi", "VII","vii")


#BLACKPINK

PlayingWithFire <-c("vi","I","IV","V",
                    "vi","I","IV","V",
                    "vi","I","IV","V",
                    "vi","V","I","vi","V",
                    "vi","I","IV","V",
                    "vi","I","IV","V",
                    "vi","V","I","vi","V",
                    "vi","I","IV","V",
                    "vi","I","IV","V",
                    "vi","I","IV","V",
                    "vi","vi","I","IV","v",
                    "vi","vi","I",
                    "vi","I","IV","V",
                    "vi","V","I","vi","V",
                    "vi","I","IV","V",
                    "vi","V","I","vi","V",
                    "vi","I","IV","V",
                    "vi","V","I","vi","V",
                    "vi","I","IV","V",
                    "vi","V","I","vi",
                    "IV", "V", "vi", "iii","III", "IV", "V", "vi","iii",
                    "IV","V","vi","V","I",
                     "IV","V","vi","V","I",
                     "IV","V","vi","V","I",
                     "IV","V") 

LovesickGirls <- c("vi", "I", "IV", "vi", "I", "IV", "V", "V", "IV", "V", "vi", "I", "IV", "V", "vi", "I", "IV", "V", "vi", "I", "IV", "V", "vi", "vi", "V", "I", "IV", "vi", "V", "I", "IV", "vi", "V", "I", "IV", "vi", "V", "I", "IV", "vi", "V", "I", "VI", "vi", "V", "I", "IV", "vi", "V", "I", "IV", "vi","V", "I", "IV", "IV","V","vii","vi","IV","V","vi","vi", "vi","V","I","IV","vi", "V", "I", "IV","vi","V","I","IV","vi","V","I","IV","vi","V","I","IV","vi","V","I","IV", "vi", "V", "I", "IV", "vi", "V", "I", "IV")

#Creating Matrix

Blackpink <- c(PlayingWithFire,LovesickGirls)

BP.chordcounts <- createSequenceMatrix(Blackpink,toRowProbs=FALSE,sanitize=FALSE, possibleStates = possible.chords)

BP.chordcounts

BP.chordprob <- createSequenceMatrix(Blackpink,toRowProbs=TRUE,sanitize=FALSE, possibleStates = possible.chords)

BP.chordprob

#Transition Probability Matrix 

BP.TM <- markovchainFit(Blackpink,sanitize=TRUE, possibleStates = possible.chords)

BP.TM


#FLEETWOOD MAC

Landslide <- c("I","V","vi","V","I","V","vi", "V","I","V","vi","V","I","V","vi","V","I","V","vi","V","I","V","vi","V","V", "I", "V","vi","IV","I","ii","V","I","V","vi","IV","I","ii","I","V","vi","V","I","V","vi", "V","I","V","vi","V","I","I","V","vi","V","I","V","vi","V","I","V","vi","V","V","I", "V", "vi","IV","I","ii","V","I","V","vi","IV","I","ii","I")

Dreams <- c("vi","vii","vi","vii","vi","vii","vi","vii",
            "vi","vii","vi","vii","vi","vii","vi","vii","vi","vii","vi","vii",
            "vi","vii","vi","vii","vi","vii","vi","vii","vi","vii","vi","vii",
            "vi","V","vi","V","vi","V","vi","V","vi","V","vi","V","vi","V","vi","V")

#Creating Matrix

FleetwoodMac <- c(Landslide,Dreams)

FM.chordcounts <- createSequenceMatrix(FleetwoodMac,toRowProbs = FALSE, sanitize=FALSE, possibleStates= possible.chords)

FM.chordcounts

FM.chordprob <- createSequenceMatrix(FleetwoodMac,toRowProbs=TRUE,sanitize=FALSE, possibleStates = possible.chords)

FM.chordprob

#Transition Probability Matrix

FM.TM <- markovchainFit(FleetwoodMac,sanitize=TRUE, possibleStates = possible.chords)

FM.TM




#Blackpink at a Glance: Looking at both songs separately

PWF <- markovchainFit(PlayingWithFire,sanitize=TRUE, possibleStates = possible.chords)

LG <- markovchainFit(LovesickGirls,sanitize=TRUE, possibleStates = possible.chords)

PWF
LG


L <- markovchainFit(Landslide,sanitize=TRUE, possibleStates = possible.chords)

D <- markovchainFit(Dreams,sanitize=TRUE, possibleStates = possible.chords)

L 
D


```

